---
title: "Análise de commits no Github"
author: "José Benardi de Souza Nunes"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

<br>

# Introdução

<br>

> The code and data employed here can be found at the [original repository](https://github.com/Benardi/commits_on_github). The data employed on this report is a sample of the commits made on some of the repositories on Github each day from 2016 to 2017.

<br>

***

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
library(tidyverse)
library(lubridate)
library(timeDate)
library(resample)

theme_set(theme_bw())
```

# Data Overview

<br>

```{r, warning=FALSE}
readr::read_csv(here::here("data/github-users-committing-filetypes.csv"),
                      progress = FALSE,
                      col_types = cols(
                        file_extension = col_character(),
                        month_day = col_integer(),
                        the_month = col_integer(),
                        the_year = col_integer(),
                        users = col_integer()
                        )) -> data
data %>% 
  glimpse()
```

<br>

* We will recreate a date object based on the dataframe columns and from it deduce whether the day of that observation was a weekend day or not.

```{r}
data %>%
  mutate(cronology = lubridate::ymd(paste(the_year,
                         the_month,
                         month_day)),
         isWeekend = timeDate::isWeekend(cronology)) -> data
data %>%
  sample_n(10)
```

## All file extensions

```{r}
data %>%
  group_by(file_extension, isWeekend) %>%
  summarise(popularity = sum(users)) %>%
  ggplot(aes(x=reorder(file_extension,popularity),
             y = popularity,
             fill=isWeekend)) + 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  labs(x="File extension",y="Number of users editing files")
```

* A lot of md and web related files (json, js, html) were commited on weekends. Looks like documenting and working on the front-end goes on during weekends, poor front-end developers...

<br>

```{r}
data %>%
  group_by(cronology,isWeekend) %>%
  summarise(popularity = sum(users)) %>%
  ggplot(aes(popularity,cronology,color=isWeekend)) +
  geom_segment(aes(x = 0, y = cronology, 
                 xend = popularity, 
                 yend = cronology),
               color = "grey50",
               size=0.15) +
  geom_point() +
  coord_flip() +
  labs(y="Point in Time",
       x="Number of users editing files")
```

* We can see a sizable decreae in user's commits around January, this matches typical holydays.

<br>

## Python and Java

```{r}
data %>%
  filter(file_extension %in% c("py","java")) %>%
  group_by(file_extension, isWeekend) %>%
  summarise(popularity = median(users)) %>%
  ggplot(aes(x=reorder(file_extension,popularity),
             y = popularity,
             fill=isWeekend)) + 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  labs(x="File extension",y="Median number of users editing files")
```

* In terms of **sample** Python seems to have more commits than Java on weekends regarding median.
* In terms of **sample** both Java and Python seem to have more files edited over weekdays than over weekends, regarding median. 

```{r}
data %>%
  filter(file_extension %in% c("py","java")) %>%
  ggplot(aes(users,cronology,color=isWeekend)) +
  geom_segment(aes(x = 0, y = cronology, 
                 xend = users, 
                 yend = cronology),
               color = "grey50",
               size=0.15) +
  geom_point() +
  facet_wrap(~ file_extension,
             nrow = 2) + 
  coord_flip() +
  labs(y="Point in Time",
       x="Number of users editing files")
```

* Both Python and Java reflect the same drop on file editions around January.

<br>

***

<br>

# Statistical Inference

<br>

> As talking about the **sample** isn't enough to draw conclusions about the population (coders in Github), further into this report we will make use of **statistical inference**.

<br>

## Java

<br>

### Overview

```{r}
data %>%
  filter(file_extension == "java") %>%
  ggplot(aes(x="", 
             y=users,
             group=isWeekend, 
             fill=isWeekend)) + 
  geom_boxplot() +
  labs(y="Number of users editing files") +
  ggtitle("Java popularity (weekdays vs weekend)") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

* In terms of **sample** we see more clearly that Java coders work way more during the week.

<br>

### Inference on two samples

* We will make use of confidence intervals at a 95% degree of confidence

```{r, results='hold'}
data %>%
  filter(file_extension == "java") %>%
  filter(!isWeekend) %>%
  bootstrap(median(users), R = 10000) %>% 
    CI.percentile(probs = c(.025, .975)) -> java.week
data %>%
  filter(file_extension == "java") %>%
  filter(isWeekend) %>%
  bootstrap(median(users), R = 10000) %>% 
    CI.percentile(probs = c(.025, .975)) -> java.weekend

cat(paste("Java on week days:\n"))
java.week
cat(paste("\n\nJava on weekend days:\n"))
java.weekend
```

<br>
<br>

```{r, results='hold'}
df = data.frame(rbind(java.week, 
                      java.weekend[rownames(java.weekend),]))

df$medida = c("JAVA (Week)", "JAVA (Weekend)")

df %>% 
  ggplot(aes(x = medida, ymin = X2.5., ymax = X97.5.)) + 
  geom_errorbar(width = .2) +
  labs(y= "Number of users editing files", x="Group")
```

<br>

>Looking at the confidence intervals (C.I.) for Java coders editions during the week and Java coders editions during the weekend **we can say at a 95% degree of confidence that there's a statistically significant difference between the median of files edited by Java coders during the week and the median of files edited by Java coders during the weekend.**  

<br>

### Inference on the unpaired difference of two samples

<br>

* Let's bootstrap the unpaired difference between the median number of java users on weekdays and the median number of java users on weekends.

<br>

```{r, results='hold'}
data %>%
  filter(file_extension == "java") -> java

b.diff.means <- bootstrap2(java$users, 
                          treatment = java$isWeekend, 
                          median, R = 10000)

means.diff = CI.percentile(b.diff.means, probs = c(.025, .975))
means.diff

data.frame(means.diff) %>%
  ggplot(aes(x = "Difference",ymin = X2.5., ymax = X97.5.)) +
  geom_errorbar(width = .2) +
  geom_hline(yintercept = 0, colour = "darkorange") +
  labs(x="")

```

* The C.I shows us that Java coders work way more during the week (Interval is exclusively above 0), that was expected given the business feel the community around Java has.

<br>

> Looking at the confidence intervals (C.I.) of the unpaired difference between the median number of java users on weekdays and the median number of java users on weekends **we can say at a 95% degree of confidence that the median of files edited by Java coders during the week is significantly bigger than the median of files edited by Java coders during the weekend**.  

<br>

***

## Python

<br>

### Overview

```{r}
data %>%
  filter(file_extension == "py") %>%
  ggplot(aes(x="", 
             y=users,
             group=isWeekend, 
             fill=isWeekend)) + 
  geom_boxplot() + 
  labs(x="", y="Number of users editing files") +
  ggtitle("Python popularity weekday vs weekend") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

* In terms of **sample** we see more clearly that Python coders work way more during the week.

<br>

### Inference on two samples

```{r, results='hold'}
data %>%
  filter(file_extension == "py") %>%
  filter(!isWeekend) %>%
  bootstrap(median(users), R = 10000) %>% 
    CI.percentile(probs = c(.025, .975)) -> python.week
data %>%
  filter(file_extension == "py") %>%
  filter(isWeekend) %>%
  bootstrap(median(users), R = 10000) %>% 
    CI.percentile(probs = c(.025, .975)) -> python.weekend

cat(paste("Python on week days:\n"))
python.week
cat(paste("\n\nPython on weekend days:\n"))
python.weekend
```

<br>
<br>

```{r}
df = data.frame(rbind(python.week, 
                      python.weekend[rownames(python.week),]))
df$medida = c("python.week", "python.weekend")

df %>% 
  ggplot(aes(x = medida, ymin = X2.5., ymax = X97.5.)) + 
  geom_errorbar(width = .2) +
  labs(y= "Number of users editing files", x="Group")
```

<br>

>Looking at the confidence intervals (C.I.) for Python coders editions during the week and Python coders editions during the weekend **we can say at a 95% degree of confidence that there's a statistically significant difference between the median of files edited by Python coders during the week and the median of files edited by Python coders during the weekend.**  

<br>

### Inference on the unpaired difference of two samples

<br>

* Let's bootstrap the unpaired difference between the median number of python users on weekdays and the median number of python users on weekends.

<br>

```{r, results='hold'}
data %>%
  filter(file_extension == "py") -> python

b.diff.means <- bootstrap2(python$users, 
                          treatment = python$isWeekend, 
                          median, R = 10000)

means.diff = CI.percentile(b.diff.means, probs = c(.025, .975))
means.diff

data.frame(means.diff) %>%
  ggplot(aes(x = "Difference",ymin = X2.5., ymax = X97.5.)) +
  geom_errorbar(width = .2) +
  geom_hline(yintercept = 0, colour = "darkorange") +
  labs(x="")
```

* The C.I shows us that Python coders work way more during the week (Interval is exclusively above 0). 
  * The former was not unexpected to be honest, given that there's a whole business segment around Python despite its hypster vibes.

<br>

> Looking at the confidence intervals (C.I.) of the unpaired difference between the median number of Python users on weekdays and the median number of Python users on weekends **we can say at a 95% degree of confidence that the median of files edited by Python coders during the week is significantly bigger than the median of files edited by Python coders during the weekend**.  

<br>

***

## Python/Java on the weekend

<br>

### Overview

```{r}
data %>%
  filter(isWeekend) %>%
  filter(file_extension %in% c("py","java")) %>%
  ggplot(aes(x=file_extension, 
             y=users,
             group=file_extension, 
             fill=file_extension)) + 
  geom_boxplot() +
  ggtitle("Python vs Java (Weekends)") +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  labs(y="Number of users editing files")
```

### Inference on two samples

```{r, results='hold'}
cat(paste("Java on weekend days:\n"))
java.weekend
cat(paste("\n\nPython on weekend days:\n"))
python.weekend
```

```{r}
df = data.frame(rbind(java.weekend, 
                      python.weekend[rownames(python.week),]))
df$medida = c("java.weekend", "python.weekend")

df %>% 
  ggplot(aes(x = medida, ymin = X2.5., ymax = X97.5.)) + 
  geom_errorbar(width = .2) +
  labs(y= "Number of users editing files", x="Group")
```

### Inference on the unpaired difference of two samples

<br>

* Let's bootstrap the unpaired difference between the median number of java and python users on weekends.

<br>

```{r, results='hold'}
data %>%
  filter(isWeekend) %>%
  filter(file_extension %in% c("py","java")) -> weekend

b.diff.means <- bootstrap2(weekend$users, 
                          treatment = weekend$file_extension, 
                          median, R = 10000)

means.diff = CI.percentile(b.diff.means, probs = c(.025, .975))
means.diff

data.frame(means.diff) %>%
  ggplot(aes(x = "Difference",ymin = X2.5., ymax = X97.5.)) +
  geom_errorbar(width = .2) +
  geom_hline(yintercept = 0, colour = "darkorange") +
  labs(x="")
```
